---
title: "Global and regional temp model"
author: "Eddie Wu"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This R markdown file is used to compare different water temperature models on a weekly timescale.

1. Weekly water temperature output from futureStream database
2. lag 5 days water temperature model - converted to weakly scale
3. ...


```{r library, warning=FALSE, message=FALSE}
library(dplyr)
library(caret)
library(tidyverse)
library(lme4)
library(nlme)
library(xts)
library(ModelMetrics)
library(zoo)
library(lubridate)
```

```{r functions}
form.fixed <- waterT ~ airT.lag1 + airT.lag2 + airT.lag3 + airT.lag4 + airT.lag5

form.locrandom <- waterT ~ airT.lag1 + airT.lag2 + airT.lag3 + airT.lag4 + airT.lag5 + (1|location)


calculate.rmse <- function(fold, training, testing, form, type) {
  
  # lists to store all the outputs
  model.list <- vector("list", fold)
  out.list <- vector("list", fold)
  rmse.list <- matrix(NA, nrow = fold,
                      ncol = length(unique(training[[1]]$location)))
  colnames(rmse.list) <- unique(training[[1]]$location)
  
  # loop starts here
  for (i in 1:fold) {
    
    compare <- NA
    
    # select current dataset, and all unique location levels
    current.training <- training[[i]]
    current.testing <- testing[[i]]
    loc.levels <- unique(current.training$location)
    
    if (type == "fixed") { # fixed effect model
      model <- lm(form, data = current.training)
      preds <- predict(model, newdata = current.testing)}
    else{ # mixed effect model
      model <- lmer(form, data = current.training)
      preds <- predict(model, newdata = current.testing, re.form=(~1|location))}
    
    p <- as.data.frame(preds)
    
    # calculate RMSE
    compare <- cbind(current.testing, preds = p$preds) %>% 
      select(location, date, obs = waterT, preds)
    
    for (loc in 1:length(loc.levels)) {
      compare.now <- compare[compare$location == loc.levels[loc],]
      rmse <- rmse(compare.now$obs, compare.now$preds)
      rmse.list[i,loc] <- rmse
    }
    
    model.list[[i]] <- model
    out.list[[i]] <- compare
  }
  return(list(rmse = rmse.list, model = model.list, out = out.list))
}
```

## Data importing and cleaning

```{r data import}
## Data import
airtemp <- read.csv("tributary air temperatures clean.csv")
watertemp <- read.csv("tributary water temperature/water_temperature_d.csv")

# Convert to date format
airtemp$date <- as.Date(airtemp$date, format = "%m/%d/%Y")
watertemp$date <- as.Date(watertemp$date, format = "%m/%d/%Y")

table(airtemp$station_name)
table(watertemp$station_name)
table(watertemp$location)


## Calculate the MEAN airT for US locations
for (i in which(is.na(airtemp$mean_temp))) {
  airtemp$mean_temp <- (airtemp$max_temp + airtemp$min_temp) / 2
}
```


### Check for imputed values

Use a 7-days rolling mean to check for possibly imputed temperatures. If the variance of a certain day's temperature is very close to zero, then it is likely that this particular data is imputed.

```{r impute check}
# make sure that no initial NA values in watertemp
which(is.na(watertemp$location))

# Need to calculate the rolling mean for each location separately...
watertemp$location <- as.factor(watertemp$location)
unique_locations <- unique(watertemp$location)

for(loc in unique_locations) {
  sub <- watertemp[watertemp$location == loc,]
  sub$rolling_mean <- rollmean(sub$temp, k = 7, fill = NA, align = "right")
  watertemp[watertemp$location == loc, "rolling_mean"] <- sub$rolling_mean
}

# Calculate variance
watertemp$variance <- (watertemp$temp - watertemp$rolling_mean)^2

# Assign NA when variance is very small
watertemp$temp[which(watertemp$variance < 1e-10)] <- NA
```

### Get lagged days

```{r lagged days}
## Data cleaning
master.temp <- left_join(airtemp, watertemp, by = c("station_name", "date")) %>% 
  select(location, country, station_name, station_id, date,
         year, month = month.x, day = day.x, airT = mean_temp, waterT = temp) %>%
  na.omit() %>% 
  arrange(location, date) # arrange by location and date


## Get the time lag day variables
master.temp <- master.temp %>%
  group_by(location) %>%
  mutate(airT.lag1 = lag(airT, 1),
         airT.lag2 = lag(airT, 2),
         airT.lag3 = lag(airT, 3),
         airT.lag4 = lag(airT, 4),
         airT.lag5 = lag(airT, 5)) %>% 
  na.omit()

# Change the location into factors
master.temp$location <- as.factor(master.temp$location)
table(master.temp$location)
```

### futureStream database

We import weekly modelled water temperature data from the 14 tributary locations.

```{r futureStream data import}
## Import all 14 files
bigcreek <-read.csv("weekly modelled water temperature/bigcreek_model.csv") %>%
  mutate(X = "bigcreek")
bigotter <- read.csv("weekly modelled water temperature/bigotter_model.csv") %>%
  mutate(X = "bigotter")
fox <- read.csv("weekly modelled water temperature/fox_model.csv") %>% 
  mutate(X = "fox")
genesee <- read.csv("weekly modelled water temperature/genesee_model.csv") %>% 
  mutate(X = "genesee")
humber <- read.csv("weekly modelled water temperature/humber_model.csv") %>% 
  mutate(X = "humber")
longpoint <- read.csv("weekly modelled water temperature/lp_model.csv") %>% 
  mutate(X = "longpoint")
mississagi <- read.csv("weekly modelled water temperature/mississagi_model.csv") %>% 
  mutate(X = "mississagi")
nipigon <- read.csv("weekly modelled water temperature/nipigon_model.csv") %>%
  mutate(X = "nipigon")
portage <- read.csv("weekly modelled water temperature/pb_model.csv") %>% 
  mutate(X = "portage")
portdover <- read.csv("weekly modelled water temperature/portdover_model.csv") %>% 
  mutate(X = "portdover")
saginaw <- read.csv("weekly modelled water temperature/saginaw_model.csv") %>%
  mutate(X = "saginaw")
still <- read.csv("weekly modelled water temperature/still_model.csv") %>%
  mutate(X = "still")
stlouis <- read.csv("weekly modelled water temperature/st_louis_model.csv") %>%
  mutate(X = "stlouis")
vermilion <- read.csv("weekly modelled water temperature/vermilion_model.csv") %>%
  mutate(X = "vermilion")

# Combine into one dataframe
futurestream.temp <- rbind(bigcreek,bigotter,fox,genesee,humber,longpoint,
                           mississagi,nipigon,portage,portdover,saginaw,
                           still,stlouis,vermilion) %>% 
  rename(location = X, week = weeks, preds.futureS = temperature.avg) %>% 
  arrange(location)
```



## SECTION 1: Lag5 day multiple linear regression

### Get training and testing

First we need to get the training and testing data for 10 fold. Notice that if we want to compare between models, we need to use the same training and testing dataset for fitting all models and calculating RMSE.

```{r get training and testing}
master.temp$country <- as.factor(master.temp$country)

## Subsetting
df_by_location <- split(master.temp, master.temp$location)
df_by_location <- df_by_location[sapply(df_by_location, function(x) !is.null(x) && nrow(x) > 0)]

fold = 10 #run 10 folds
combined_training_list <- vector("list",fold)
combined_testing_list <- vector("list",fold)

## Loop
for (f in 1:fold) {
  
  for (loc in 1:length(df_by_location)) {
    
    # Subset the current location data, and get train/test
    current <- df_by_location[[loc]]
    current_training <- current[current$year == sample(current$year, 1),]
    current_testing <- current[current$year == sample(current$year, 1),]
    
    # Add to the combined list
    combined_training_list[[f]] <- rbind(
      combined_training_list[[f]], current_training)
    combined_testing_list[[f]] <- rbind(
      combined_testing_list[[f]], current_testing)
  }
  
  combined_training_list[[f]]$year <- as.factor(combined_training_list[[f]]$year)
  combined_testing_list[[f]]$year <- as.factor(combined_testing_list[[f]]$year)
}
```

### Linear lag 5 days multiple regression model (both fixed and mixed).

```{r lag5 linear}
## Get results using 10-fold iterations
results.fixed <- calculate.rmse(fold,combined_training_list,
                                combined_testing_list,
                                form.fixed,"fixed")
results.mixed <- calculate.rmse(fold, combined_training_list,
                               combined_testing_list,
                               form.locrandom,"mixed")

# Get RMSE
rmse.lag5.f <- results.fixed[["rmse"]]
rmse.lag5.m <- results.mixed[["rmse"]]

colMeans(rmse.lag5.f)
colMeans(rmse.lag5.m)
```



## SECTION 2: Compare lag5 and futureS on a weekly time scale

1. Convert the lag5 model daily water temperature predictions to weekly water temperature predictions.

2. Merge the weekly observations with predictions from lag5 model and predictions from futureStream model.

3. Calculate RMSE for both models and compare (for each location).

*Note: Each is done through 10 iterations.*

```{r convert lag5 modelled output to weekly and compare to futureS}
# Dataframe to store the results
rmse.lag5 <- matrix(NA, nrow = 10, ncol = length(unique_locations))
colnames(rmse.lag5) <- unique_locations

rmse.futureS <- matrix(NA, nrow = 10, ncol = length(unique_locations))
colnames(rmse.futureS) <- unique_locations

## Loop through 10 folds
for(i in 1:fold) {
  ## Convert lag5 model prediction to weekly timescale
  df.days <- results.mixed[["out"]][[i]]
  df.days$week <- week(df.days$date)
  df.days$week[df.days$week == 53] <- 52 #need to convert week 53 to 52
  
  df.week <- merge(
    aggregate(preds~week+location, FUN=mean, data=df.days, na.rm=TRUE),
    aggregate(obs~week+location, FUN=mean, data=df.days, na.rm=TRUE),
    all=TRUE)
  
  # Merge together
  temp.compare <- merge(df.week, futurestream.temp) %>% 
    rename(preds.lag5 = preds) %>% 
    select(week, location, obs, preds.lag5, preds.futureS) %>% 
    arrange(location)
  
  # get RMSE for each location
  for (loc in 1:length(unique_locations)) {
    compare <- temp.compare[temp.compare$location == unique_locations[loc],]
    rmse.lag5[i,loc] <- rmse(compare$obs, compare$preds.lag5)
    rmse.futureS[i,loc] <- rmse(compare$obs, compare$preds.futureS)
  }
}

colMeans(rmse.lag5)
colMeans(rmse.futureS)
```



## SECTION 3: Stochastic model

Simple stochastic model with a annual component and a short-term residual component on the daily water and air temperature data. Model formulation is:

$$T_W(t) = T_A(t) + R_W(t)$$

Annual component is calculated by optimizing a sine function to the times series.

$$T_A(t) = a + b*sin[\frac{2\pi}{365}(t+t_0)]$$
, where a, b, and t0 are estimated coefficients.


Short-term residual component is based on multiple regression lagged air temperature residuals (based on Kothandaraman, 1971). (Note that there are other ways of representing the residual components...)

$$R_W(t) = \beta_1R_A(t) + \beta_2R_A(t-1) + \beta_3R_A(t-2)$$
, where $\beta_1$, $\beta_2$, $\beta_3$ are regression coefficients, and $R_A(t)$, $R_A(t-1)$, $R_A(t-2)$ are air temperature residuals at time t, t-1, and t-2.


### AirT residuals as short-term component

We first try a stochastic model with no random effects.

```{r stochastic fixed effect}
out <- vector("list", fold)
rmse <- matrix(NA, nrow=fold,
               ncol=length(unique(combined_training_list[[1]]$location)))
colnames(rmse) <- unique(combined_training_list[[1]]$location)

# Use a list to store all the outputs
results.sto.f <- list(rmse = rmse, out = out)


## Iteration starts here
for (i in 1:fold){
  
  compare <- NA
  # select current dataset, and all unique location levels
  current.training <- combined_training_list[[i]]
  current.testing <- combined_testing_list[[i]]
  loc.levels <- unique(current.training$location)

  ## TRAINING
  # get the model annual component
  annual.comp <- nls(airT ~ a+b*sin(2*pi/365*(yday(date)+t0)),
                    start = list(a=0.05, b=5, t0=-26), data=current.training)
  current.training$airT - fitted(annual.comp)
  
  # get the air temperature residuals
  res <- as.data.frame(matrix(NA, ncol = 4, 
                              nrow = length(current.training$airT)))
                              # dataframe to store the residuals
  colnames(res) <- c("res.t", "res.t1", "res.t2", "res.w")
  res[,"res.t"] <- as.vector(residuals(annual.comp))
  res[,"res.t1"][-1] <- res[,"res.t"][-nrow(res)]
  res[,"res.t2"][-1] <- res[,"res.t1"][-nrow(res)]
  res[,"res.w"] <- residuals(nls(waterT ~ a+b*sin(2*pi/365*(yday(date)+t0)),
                   start = list(a=0.05, b=5, t0=-26),
                   data = current.training))
  
  # get the water temperature residual component
  residual.comp <- lm(res.w ~ res.t + res.t1 + res.t2,
                      data = res, na.action = na.omit)
  
  ## TESTING
  preds.annual <- predict(annual.comp, newdata=current.testing)

  res <- as.data.frame(matrix(NA, ncol = 3, 
                            nrow = length(current.testing$airT))) #residuals
  colnames(res) <- c("res.t", "res.t1", "res.t2")
  res[,"res.t"] <- current.testing$airT - preds.annual
  res[,"res.t1"][-1] <- res[,"res.t"][-nrow(res)]
  res[,"res.t2"][-1] <- res[,"res.t1"][-nrow(res)]
  preds.residuals <- predict(residual.comp, newdata=res)
  
  # add up both components
  p <- as.data.frame(preds.annual + preds.residuals)

  ## Calculate RMSE
  compare <- cbind(current.testing,
                   preds = p$`preds.annual + preds.residuals`) %>% 
  select(location, date, obs = waterT, preds)

  for (loc in 1:length(loc.levels)) {
    compare.now <- compare[compare$location == loc.levels[loc],] %>% na.omit()
    results.sto.f[["rmse"]][i,loc] <- 
      rmse(compare.now$obs, compare.now$preds)
  }
  
  results.sto.f[["out"]][[i]] <- compare
}


## Get the results
rmse.sto.f <- results.sto.f[["rmse"]]
colMeans(rmse.sto.f)
```

Then, we try a stochastic model with mixed effect (location as the random factor).

```{r stochastic mixed effect}
out <- vector("list", fold)
rmse <- matrix(NA, nrow=fold,
               ncol=length(unique(combined_training_list[[1]]$location)))
colnames(rmse) <- unique(combined_training_list[[1]]$location)

# Use a list to store all the outputs
results.sto.m <- list(rmse = rmse, out = out)


for (i in 1:fold){
  
  compare <- NA
  # select current dataset, and all unique location levels
  current.training <- combined_training_list[[i]]
  current.testing <- combined_testing_list[[i]]
  loc.levels <- unique(current.training$location)
  
  ## TRAINING
  # get the model annual component
  annual.comp <- nls(airT ~ a+b*sin(2*pi/365*(yday(date)+t0)),
                    start = list(a=0.05, b=5, t0=-26), data=current.training)

  # get the air temperature residuals
  res <- as.data.frame(matrix(NA, ncol = 5, 
                              nrow = length(current.training$airT)))
                              # dataframe to store the residuals
  colnames(res) <- c("res.t", "res.t1", "res.t2", "res.w", "location")
  res[,"res.t"] <- as.vector(residuals(annual.comp))
  res[,"res.t1"][-1] <- res[,"res.t"][-nrow(res)]
  res[,"res.t2"][-1] <- res[,"res.t1"][-nrow(res)]
  res[,"res.w"] <- residuals(nls(waterT ~ a+b*sin(2*pi/365*(yday(date)+t0)),
                   start = list(a=0.05, b=5, t0=-26),
                   data = current.training))
  res[,"location"] <- current.training$location

  # get the water temperature residual component
  residual.comp <- lmer(res.w ~ res.t + res.t1 + res.t2 + (1|location),
                       data = res, na.action = na.omit)
  
  ## TESTING
  preds.annual <- predict(annual.comp, newdata=current.testing)

  res <- as.data.frame(matrix(NA, ncol = 4, 
                            nrow = length(current.testing$airT))) #residuals
  colnames(res) <- c("res.t", "res.t1", "res.t2", "location")
  res[,"res.t"] <- current.testing$airT - preds.annual
  res[,"res.t1"][-1] <- res[,"res.t"][-nrow(res)]
  res[,"res.t2"][-1] <- res[,"res.t1"][-nrow(res)]
  res[,"location"] <- current.testing$location
  preds.residuals <- predict(residual.comp, newdata=res,
                             re.form=(~1|location))
  
  # add up both components
  p <- as.data.frame(preds.annual + preds.residuals)

  ## Calculate RMSE
  compare <- cbind(current.testing,
                   preds = p$`preds.annual + preds.residuals`) %>% 
  select(location, date, obs = waterT, preds)

  for (loc in 1:length(loc.levels)) {
    compare.now <- compare[compare$location == loc.levels[loc],] %>% na.omit()
    results.sto.m[["rmse"]][i,loc] <- rmse(compare.now$obs, compare.now$preds)
  }
    
  results.sto.m[["out"]][[i]] <- compare
}


## Get the results
rmse.sto.m <- results.sto.m[["rmse"]]
colMeans(rmse.sto.m)
```

### Actual airT as short-term component

Some literatures (Rabi et al., 2015) use actual air temperatures (instead of residuals) to represent the residual component. In this case, the residual component is:

$$R_W(t) = \beta_1T_A(t) + \beta_2T_A(t-1) + \beta_3T_A(t-2)$$

We first start with a fixed model.

```{r stochastic fixed actual air}
out <- vector("list", fold)
rmse <- matrix(NA, nrow=fold,
               ncol=length(unique(combined_training_list[[1]]$location)))
colnames(rmse) <- unique(combined_training_list[[1]]$location)

# Use a list to store all the outputs
results.sto.fa <- list(rmse = rmse, out = out)


## Iteration starts here
for (i in 1:fold){
  
  compare <- NA
  # select current dataset, and all unique location levels
  current.training <- combined_training_list[[i]]
  current.testing <- combined_testing_list[[i]]
  loc.levels <- unique(current.training$location)

  ## TRAINING
  # get the model annual component
  annual.comp <- nls(airT ~ a+b*sin(2*pi/365*(yday(date)+t0)),
                    start = list(a=0.05, b=5, t0=-26), data=current.training)
  
  # get the water temperature residual component
  residual.comp <- lm(waterT ~ airT + airT.lag1 + airT.lag2,
                      data = current.training)
  
  ## TESTING
  preds.annual <- predict(annual.comp, newdata=current.testing)
  preds.residuals <- predict(residual.comp, newdata=current.testing)
  p <- as.data.frame(preds.annual + preds.residuals)

  ## Calculate RMSE
  compare <- cbind(current.testing,
                   preds = p$`preds.annual + preds.residuals`) %>% 
  select(location, date, obs = waterT, preds)

  for (loc in 1:length(loc.levels)) {
    compare.now <- compare[compare$location == loc.levels[loc],] %>% na.omit()
    results.sto.fa[["rmse"]][i,loc] <- 
      rmse(compare.now$obs, compare.now$preds)
  }
  
  results.sto.fa[["out"]][[i]] <- compare
}


## Get the results
rmse.sto.fa <- results.sto.fa[["rmse"]]
colMeans(rmse.sto.fa)
```

Now a mixed model with location as a random effect

```{r stochastic mixed actual air}
out <- vector("list", fold)
rmse <- matrix(NA, nrow=fold,
               ncol=length(unique(combined_training_list[[1]]$location)))
colnames(rmse) <- unique(combined_training_list[[1]]$location)

# Use a list to store all the outputs
results.sto.ma <- list(rmse = rmse, out = out)


## Iteration starts here
for (i in 1:fold){
  
  compare <- NA
  # select current dataset, and all unique location levels
  current.training <- combined_training_list[[i]]
  current.testing <- combined_testing_list[[i]]
  loc.levels <- unique(current.training$location)

  ## TRAINING
  # get the model annual component
  annual.comp <- nls(airT ~ a+b*sin(2*pi/365*(yday(date)+t0)),
                    start = list(a=0.05, b=5, t0=-26), data=current.training)
  
  # get the water temperature residual component
  residual.comp <- lmer(waterT ~ airT + airT.lag1 + airT.lag2 + (1|location),
                        data = current.training)
  
  ## TESTING
  preds.annual <- predict(annual.comp, newdata=current.testing)
  preds.residuals <- predict(residual.comp, newdata=current.testing,
                             re.form=(~1|location))
  p <- as.data.frame(preds.annual + preds.residuals)

  ## Calculate RMSE
  compare <- cbind(current.testing,
                   preds = p$`preds.annual + preds.residuals`) %>% 
  select(location, date, obs = waterT, preds)

  for (loc in 1:length(loc.levels)) {
    compare.now <- compare[compare$location == loc.levels[loc],] %>% na.omit()
    results.sto.ma[["rmse"]][i,loc] <- 
      rmse(compare.now$obs, compare.now$preds)
  }
  
  results.sto.ma[["out"]][[i]] <- compare
}


## Get the results
rmse.sto.ma <- results.sto.ma[["rmse"]]
colMeans(rmse.sto.ma)
```

Both models with actual air temperature seem to perform very poorly... So we should not use this form...


## SECTION 4: Overall Comparison and graphs
```{r rmse comparison table, eval = FALSE}
rmse.lag5.f
rmse.lag5.m
rmse.sto.f
rmse.sto.m

# regional comparison
data.frame(lag5_fixed = colMeans(rmse.lag5.f),
           lag5_mixed = colMeans(rmse.lag5.m),
           sto_fixed = colMeans(rmse.sto.f),
           sto_mixed = colMeans(rmse.sto.m))

# global and regional
a <- data.frame(regional_lag5 = colMeans(rmse.lag5),
                global = colMeans(rmse.futureS))

View(a)

max(a[,"regional_lag5"])
max(a[,"global"])

```


```{r plot, eval = FALSE}
a <- coef(annual.comp)[1]
b <- coef(annual.comp)[2]
t0 <- coef(annual.comp)[3]

form <- function(x) a+b*sin(2*pi/365*(x+t0))

curve(expr = form, from=1, to=1000)


pp.c <- compare %>% filter(location == "mississagi")

ggplot(data=pp.c, aes(x=date))+
  geom_line(aes(x=date, y=obs), color = "black")+
  geom_line(aes(x=date, y=preds), color = "red")


```
