---
title: "Global and regional temperature model comparison"
author: "Eddie Wu"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This R markdown file is used to 

1. Import and clean global water temperature model

2. Compare the weekly performance of a non-linear model and the global futureStreams river temperature model on a different seasonal scales (spring, summer, fall, annual).


```{r library, warning=FALSE, message=FALSE}
library(dplyr)
library(tidyverse)
library(lme4)
library(nlme)
library(xts)
library(ModelMetrics)
library(zoo)
library(lubridate)
library(nls2)
```

```{r functions}
get.traintest <- function(master.df, year.list, fold) {
  
  # create output
  combined_training_list <- vector("list",fold)
  combined_testing_list <- vector("list",fold)
  
  ## Loop 10 folds
  for (f in 1:fold) {
    
    # dataframe to store the sampled years for each location
    dfind = data.frame(location=character(),
                     year=integer(), stringsAsFactors = FALSE)
    dfindt = data.frame(location=character(),
                     year=integer(), stringsAsFactors = FALSE)
    
    for (i in 1:length(year.list)){ # i loops through each location
      smp = year.list[[i]] #train
      if (length(smp)>1){
        indx.train=(sample(smp, 1))
        ytrain=indx.train
      } else if (length(smp)==1){
        indx.train=smp
        ytrain=indx.train
      }
      
      smpt = year.list[[i]][year.list[[i]] != indx.train] #test
      if (length(smpt)>1){
        indx.test=(sample(smpt, 1))
        ytest=indx.test
      } else if (length(smpt)==1){
        indx.test=smpt
        ytest=indx.test
      }
      
      dfind[i,] = c(names(year.list[i]), ytrain)
      dfindt[i,] = c(names(year.list[i]), ytest)
    }
    
    # subset the dataframe
    awf_train <- merge(master.df, dfind, by=c("location", "year"))
    awf_test <- merge(master.df, dfindt, by=c("location", "year"))
      
    combined_training_list[[f]] <- awf_train
    combined_testing_list[[f]] <- awf_test
  }
  return(list(combined_training_list, combined_testing_list))
}
good_year <- function(master.df, dayln) {
  # get table of sample years by location
  yr_out=table(master.df$location, master.df$year)

  # select sample years with more than X days (x=250)
  full_year=list()
  sind=vector()
  cnt=0
  ind=0
  
  for (i in seq_along(loc_seq)){
   rm(ind)
   if (ncol(yr_out)>=1) {
     ind=which(yr_out[row.names(yr_out)==loc_seq[i],]>dayln)
     
     if (length(ind)>0) {
       sind=c(sind,i)
       cnt=cnt+1
       full_year[[cnt]]=colnames(yr_out)[ind]
     }
   }
  }
  full_year=setNames(full_year,loc_seq[sind])
  print(full_year)
}
```



## Data importing and cleaning

```{r data import}
## Data import
air <- read.csv("tributary air temperatures clean.csv", stringsAsFactors=F)
water <- read.csv("tributary water temperature/water_temperature_d.csv",
                  stringsAsFactors=F)
flow <- read.csv("tributary discharge.csv", stringsAsFactors=F)

# Convert to date format
air$date <- as.Date(air$date, "%m/%d/%Y")
water$date <- as.Date(water$date, "%m/%d/%Y")
flow$date <- as.Date(flow$date, "%m/%d/%Y")


## Calculate the MEAN airT for US locations
for (i in which(is.na(air$mean_temp))) {
  air$mean_temp <- (air$max_temp + air$min_temp) / 2
}
```

### Data combining

```{r data combine}
## Combine water temperature and discharge
aw <- merge(air, water, by = c("station_name","date"))
awf <- merge(aw, flow, by = c("location","date"))
awf <- awf %>% arrange(location, date)

## Change into factor
awf$location <- as.factor(awf$location)
awf$station_name <- as.factor(awf$station_name)

# Get location sequence
loc_seq=levels(awf$location)


## Check if there are any duplicates
duplicates <- awf %>%
  group_by(location, date) %>%
  filter(n() > 1) # Duplicates should be NA...


## Print the result
table(awf$location)
```

### Check for imputed values

Use a 7-days rolling mean to check for possibly imputed temperatures. If the variance of a certain day's temperature is very close to zero, then it is likely that this particular data is imputed.

```{r impute check}
# make sure that no initial NA values in awf water temperature
which(is.na(awf$temp))

# Need to calculate the rolling mean for each location separately...
for(loc in loc_seq) {
  sub <- awf[awf$location == loc,]
  sub$rolling_mean <- rollmean(sub$temp, k = 7, fill = NA, align = "right")
  awf[awf$location == loc, "rolling_mean"] <- sub$rolling_mean
}

# Calculate variance
awf$variance <- (awf$temp - awf$rolling_mean)^2

# Assign NA when variance is very small
awf$temp[which(awf$variance < 1e-10)] <- NA

# Check results - how many imputed values are in each location
awf %>%
  group_by(location) %>%
  summarise(na_count = sum(is.na(temp)))

```


### Get lagged days

```{r lagged days}
## Get the time lag day variables
awf <- awf %>%
  group_by(location) %>%
  mutate(dmean_1 = lag(mean_temp, 1),
         dmean_2 = lag(mean_temp, 2),
         dmean_3 = lag(mean_temp, 3),
         dmean_4 = lag(mean_temp, 4),
         dmean_5 = lag(mean_temp, 5),
         dflow_1 = lag(flow, 1))


## Get cumulative air temp for past five days
awf <- awf %>% 
  rowwise() %>% 
  mutate(cair = (mean_temp+dmean_1+dmean_2+dmean_3+dmean_4+dmean_5)/6)

## Get relative flow
awf <- awf %>% mutate(rqc = (flow - dflow_1)/flow)
```


### Get final master temp dataframe

```{r master temp}
master.temp <- awf[complete.cases(cbind(awf$mean_temp,awf$temp)),] %>% 
  select(location, date, station_name, country,
         year, month = month.x, day = day.x,
         water = temp, air = mean_temp, dmean_1, dmean_2, dmean_3,
         dmean_4, dmean_5, flow, dflow_1, rqc, cair)

master.temp <- master.temp[complete.cases(master.temp[,8:18]),]
```



## Get training and testing


### Subsetting seasonal scales

Now we want to subset three master dataframes that contains seasonal-scale data. We categorize the data into four different seasonal categories:

1. spring: 3,4,5
2. summer: 6,7,8
3. fall: 9,10,11
4. winter: 12,1,2

```{r subset season}
master.sum <- master.temp %>% 
  filter(month == 6 | month == 7 | month == 8)

master.win <- master.temp %>% 
  filter(month == 12 | month == 1 | month == 2)

master.spring <- master.temp %>% 
  filter(month == 3 | month == 4 | month == 5)

master.fall <- master.temp %>% 
  filter(month == 9 | month == 10 | month == 11)

master.annual <- master.temp %>% 
  filter(month != 12 & month != 1 & month != 2)
```


### Identify good year/month data

We want the seasonal data to be greater than 60 days, annual data (winter removed) more than 180 days.

```{r good years, echo=FALSE}
## Annual (250 days)
fulyear <- good_year(master.annual, 180)


## Seasonal (60 days)
fulspring <- good_year(master.spring, 60)
fulsum <- good_year(master.sum, 60)
fulfall <- good_year(master.fall, 60)
fulwin <- good_year(master.win, 60)
```


### Get training and testing

Similarly, get training and testing for the specific season.

```{r get training and testing}
fold = 10

## Get training and testing for annual
annual <- get.traintest(master.annual, fulyear, 10)

combined_training_list <- annual[[1]]
combined_testing_list <- annual[[2]]


## Get training and testing for each season
sum <- get.traintest(master.sum, fulsum, 10)
win <- get.traintest(master.win, fulwin, 10)
spr <- get.traintest(master.spring, fulspring, 10)
fall <- get.traintest(master.fall, fulfall, 10)

combined_training_list_sp <- spr[[1]]
combined_testing_list_sp <- spr[[2]]

combined_training_list_su <- sum[[1]]
combined_testing_list_su <- sum[[2]]

combined_training_list_fa <- fall[[1]]
combined_testing_list_fa <- fall[[2]]

combined_training_list_w <- win[[1]]
combined_testing_list_w <- win[[2]]


## Create a list to store all the combined training and testing lists
grand_training <- list(combined_training_list_sp, combined_training_list_su,
                       combined_training_list_fa, combined_training_list_w,
                       combined_training_list)

grand_testing <- list(combined_testing_list_sp, combined_testing_list_su,
                       combined_testing_list_fa, combined_testing_list_w,
                       combined_testing_list)
```


## GLOBAL: futureStream database

We import weekly modelled water temperature data from the 14 tributary locations.

```{r futureStream data import}
## Import all 14 files
bigcreek <-read.csv("weekly modeled water temperature_new/bigcreek_model.csv") %>%
  mutate(X = "bigcreek")
bigotter <- read.csv("weekly modeled water temperature_new/bigotter_model.csv") %>%
  mutate(X = "bigotter")
fox <- read.csv("weekly modeled water temperature_new/fox_model.csv") %>% 
  mutate(X = "fox")
genesee <- read.csv("weekly modeled water temperature_new/genesee_model.csv") %>% 
  mutate(X = "genesee")
humber <- read.csv("weekly modeled water temperature_new/humber_model.csv") %>% 
  mutate(X = "humber")
longpoint <- read.csv("weekly modeled water temperature_new/lp_model.csv") %>% 
  mutate(X = "longpoint")
mississagi <- read.csv("weekly modeled water temperature_new/mississagi_model.csv") %>% 
  mutate(X = "mississagi")
nipigon <- read.csv("weekly modeled water temperature_new/nipigon_model.csv") %>%
  mutate(X = "nipigon")
portage <- read.csv("weekly modeled water temperature_new/pb_model.csv") %>% 
  mutate(X = "portage")
portdover <- read.csv("weekly modeled water temperature_new/portdover_model.csv") %>% 
  mutate(X = "portdover")
saginaw <- read.csv("weekly modeled water temperature_new/saginaw_model.csv") %>%
  mutate(X = "saginaw")
still <- read.csv("weekly modeled water temperature_new/still_model.csv") %>%
  mutate(X = "still")
stlouis <- read.csv("weekly modeled water temperature_new/st_louis_model.csv") %>%
  mutate(X = "stlouis")
vermilion <- read.csv("weekly modeled water temperature_new/vermilion_model.csv") %>%
  mutate(X = "vermilion")

# Combine into one dataframe
futurestream.temp <- rbind(bigcreek,bigotter,fox,genesee,humber,longpoint,
                           mississagi,nipigon,portage,portdover,saginaw,
                           still,stlouis,vermilion) %>% 
  rename(location = X, week = weeks, preds.futureS = temperature.avg) %>% 
  arrange(location)
```


## REGIONAL: non-linear model with 5 day average temperature

### Fit non-linear model
```{r non-linear, eval=TRUE}
## starting parameters
coef.spring <- c(alpha=30, gamma=0.05,beta=10)
coef.summer <- c(alpha=25, gamma=0.05,beta=10)
coef.fall <- c(alpha=30, gamma=0.05,beta=10)
coef.annual <- c(alpha=20, gamma=0.05,beta=9)

coef.list <- list(coef.spring, coef.summer, coef.fall, NA, coef.annual)


spring.r <- vector("list", fold)
summer.r <- vector("list", fold)
fall.r <- vector("list", fold)
winter.r <- vector("list", fold)
annual.r <- vector("list", fold)

grand.nonlinear <- list(spring.r, summer.r, fall.r, winter.r, annual.r)


## Iteration starts here
for (season in c(1,2,3,5)) {
  
  ## Get current season and its correspnding training/testing
  train <- grand_training[[season]]
  test <- grand_testing[[season]]
  
  ## 10 fold iteration starts here
  for (i in 1:fold){
  
    compare <- NA
    # select current dataset, and all unique location levels
    current.training <- train[[i]]
    current.testing <- test[[i]]
    compare <- current.testing

    # model training and predicting
    model <- nlme(water ~ alpha / (1 + exp(gamma * (beta - cair))),
                  fixed = list(alpha~1,gamma~1,beta~1),
                  random = gamma ~ 1|location,
                  start = coef.list[[season]],
                  data = current.training, na.action = na.omit,
                  control = list(msMaxIter = 200))
    
    compare$preds.ar <- as.vector(predict(
      model, newdata = current.testing, re.form = ~1|location))
  
    grand.nonlinear[[season]][[i]] <- compare
  }
}
```


### Convert into weekly output

```{r convert to weekly}
# new dataframe to store all weekly results
grand.week = grand.nonlinear

for (season in c(1,2,3,5)) {
  
  for(i in 1:fold) {
    ## Convert lag5 model prediction to weekly timescale
    df.days <- grand.nonlinear[[season]][[i]]
    df.days$week <- week(df.days$date)
    df.days$week[df.days$week == 53] <- 52 #need to convert week 53 to 52
  
    df.week <- merge(
      aggregate(preds.ar~week+location,FUN=mean,data=df.days,na.action=na.omit),
      aggregate(water~week+location,FUN=mean,data=df.days,na.action=na.omit),
      all=TRUE) %>% 
      na.omit() %>% 
      arrange(location, week)
  
    comp.week <- merge(df.week, futurestream.temp, by=c("location","week")) %>% 
      arrange(location, week)
  
    # store it back into the grand.nonlinear list
    grand.week[[season]][[i]] <- comp.week
  }
}
```


## Compare between GLOBAL and REGIONAL

### Calculate RMSE for each seasonal scale
```{r rmse}
# new function - different from other files...
cal.rmse <- function(out.list, fold, switch) {
  
  # data frame to store the results
  r <- matrix(NA, nrow=fold, ncol=length(unique(out.list[[1]]$location)))
  colnames(r) <- unique(out.list[[1]]$location)
  
  # 10 iterations
  for (i in 1:fold){
   compare <- out.list[[i]]
    
   # calculate rmse
   for (loc in unique(compare$location)) {
     compare.now <- subset(compare, location == loc) %>% na.omit()
     
     if (switch=="r"){
       r[i,loc] <-
         round(sqrt(mean((compare.now$water-compare.now$preds.ar)^2)),2)
     } else if (switch=="g"){
       r[i,loc] <-
         round(sqrt(mean((compare.now$water-compare.now$preds.futureS)^2)),2)
     }
    }
  }
  return(r)
}


## Annual
annual.compare <- data.frame(
  regional = colMeans(cal.rmse(grand.week[[5]], fold, "r"), na.rm=T),
  global = colMeans(cal.rmse(grand.week[[5]], fold, "g"), na.rm=T))

colMeans(annual.compare)
knitr::kable(annual.compare, digits = 3)


## Spring
spring.compare <- data.frame(
  regional = colMeans(cal.rmse(grand.week[[1]], fold, "r"), na.rm=T),
  global = colMeans(cal.rmse(grand.week[[1]], fold, "g"), na.rm=T))

colMeans(spring.compare)
knitr::kable(spring.compare, digits = 3)


## Summer
summer.compare <- data.frame(
  regional = colMeans(cal.rmse(grand.week[[2]], fold, "r"), na.rm=T),
  global = colMeans(cal.rmse(grand.week[[2]], fold, "g"), na.rm=T))

colMeans(summer.compare)
knitr::kable(summer.compare, digits = 3)


## Fall
fall.compare <- data.frame(
  regional = colMeans(cal.rmse(grand.week[[3]], fold, "r"), na.rm=T),
  global = colMeans(cal.rmse(grand.week[[3]], fold, "g"), na.rm=T))

colMeans(fall.compare)
knitr::kable(fall.compare, digits = 3)
```


### Calculate NSC for each seasonal scale
```{r nsc}
# new function - different from other files...
cal.nsc <- function(out.list, fold, switch) {
  
  # data frame to store the results
  r <- matrix(NA, nrow=fold, ncol=length(unique(out.list[[1]]$location)))
  colnames(r) <- unique(out.list[[1]]$location)
  
  # 10 iterations
  for (i in 1:fold){
   compare <- out.list[[i]]
    
   # calculate rmse
   for (loc in unique(compare$location)) {
     compare.now <- subset(compare, location == loc) %>% na.omit()
     
     if (switch=="r"){
       r[i,loc] <-
         1 - sum((compare.now$water - compare.now$preds.ar)^2) / sum((compare.now$water - mean(compare.now$water))^2)
     } else if (switch=="g"){
       r[i,loc] <-
         1 - sum((compare.now$water - compare.now$preds.futureS)^2) / sum((compare.now$water - mean(compare.now$water))^2)
     }
    }
  }
  return(r)
}


## Annual
annual.compare <- data.frame(
  regional = colMeans(cal.nsc(grand.week[[5]], fold, "r"), na.rm=T),
  global = colMeans(cal.nsc(grand.week[[5]], fold, "g"), na.rm=T))

colMeans(annual.compare)
knitr::kable(annual.compare, digits = 3)


## Spring
spring.compare <- data.frame(
  regional = colMeans(cal.nsc(grand.week[[1]], fold, "r"), na.rm=T),
  global = colMeans(cal.nsc(grand.week[[1]], fold, "g"), na.rm=T))

colMeans(spring.compare)
knitr::kable(spring.compare, digits = 3)


## Summer
summer.compare <- data.frame(
  regional = colMeans(cal.nsc(grand.week[[2]], fold, "r"), na.rm=T),
  global = colMeans(cal.nsc(grand.week[[2]], fold, "g"), na.rm=T))

colMeans(summer.compare)
knitr::kable(summer.compare, digits = 3)


## Fall
fall.compare <- data.frame(
  regional = colMeans(cal.nsc(grand.week[[3]], fold, "r"), na.rm=T),
  global = colMeans(cal.nsc(grand.week[[3]], fold, "g"), na.rm=T))

colMeans(fall.compare)
knitr::kable(fall.compare, digits = 3)
```