---
title: "Seasonal river temperature model"
author: "Eddie Wu"
date: '`r Sys.Date()`'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



## Introduction

This R markdown file is used to

1. Fit candidate models on a seasonal scale (all four seasons, and annual).

2. Get model coefficients and performance metrics calculations from all candidate models.

**All Models are mixed-effect with location as the random factor.*


```{r library, warning=FALSE, message=FALSE}
library(dplyr)
library(tidyverse)
library(lme4)
library(nlme)
library(xts)
library(ModelMetrics)
library(zoo)
library(lubridate)
library(nls2)
```

```{r functions}
cal.rmse <- function(out.list, fold) {
  
  # data frame to store the results
  r <- matrix(NA, nrow=fold, ncol=length(unique(out.list[[1]]$location)))
  colnames(r) <- unique(out.list[[1]]$location)
  
  # 10 iterations
  for (i in 1:fold){
   compare <- out.list[[i]]
    
   # calculate rmse
   for (loc in unique(compare$location)) {
     compare.now <- subset(compare, location == loc) %>% na.omit()
     r[i,loc] <- ModelMetrics::rmse(compare.now$water, compare.now$preds.ar)
    }
  }
  return(r)
}
cal.nsc <- function(out.list, fold) {
  
  # dataframe to store the results
  r <- matrix(NA, nrow=fold, ncol=length(unique(out.list[[1]]$location)))
  colnames(r) <- unique(out.list[[1]]$location)
  
  # 10 iterations
  for (i in 1:fold){
   compare <- out.list[[i]]
    
   # calculate nsc
   for (loc in unique(compare$location)) {
     compare.now <- subset(compare, location == loc) %>% na.omit()
     nsc <- 1 - sum((compare.now$water - compare.now$preds.ar)^2) / sum((compare.now$water - mean(compare.now$water))^2)
     r[i,loc] <- nsc
    }
  }
  return(r)
}
get.traintest <- function(master.df, year.list, fold) {
  
  # create output
  combined_training_list <- vector("list",fold)
  combined_testing_list <- vector("list",fold)
  
  ## Loop 10 folds
  for (f in 1:fold) {
    
    # dataframe to store the sampled years for each location
    dfind = data.frame(location=character(),
                     year=integer(), stringsAsFactors = FALSE)
    dfindt = data.frame(location=character(),
                     year=integer(), stringsAsFactors = FALSE)
    
    for (i in 1:length(year.list)){ # i loops through each location
      smp = year.list[[i]] #train
      if (length(smp)>1){
        indx.train=(sample(smp, 1))
        ytrain=indx.train
      } else if (length(smp)==1){
        indx.train=smp
        ytrain=indx.train
      }
      
      smpt = year.list[[i]][year.list[[i]] != indx.train] #test
      if (length(smpt)>1){
        indx.test=(sample(smpt, 1))
        ytest=indx.test
      } else if (length(smpt)==1){
        indx.test=smpt
        ytest=indx.test
      }
      
      dfind[i,] = c(names(year.list[i]), ytrain)
      dfindt[i,] = c(names(year.list[i]), ytest)
    }
    
    # subset the dataframe
    awf_train <- merge(master.df, dfind, by=c("location", "year"))
    awf_test <- merge(master.df, dfindt, by=c("location", "year"))
      
    combined_training_list[[f]] <- awf_train
    combined_testing_list[[f]] <- awf_test
  }
  return(list(combined_training_list, combined_testing_list))
}
good_year <- function(master.df, dayln) {
  # get table of sample years by location
  yr_out=table(master.df$location, master.df$year)

  # select sample years with more than X days (x=250)
  full_year=list()
  sind=vector()
  cnt=0
  ind=0
  
  for (i in seq_along(loc_seq)){
   rm(ind)
   if (ncol(yr_out)>=1) {
     ind=which(yr_out[row.names(yr_out)==loc_seq[i],]>dayln)
     
     if (length(ind)>0) {
       sind=c(sind,i)
       cnt=cnt+1
       full_year[[cnt]]=colnames(yr_out)[ind]
     }
   }
  }
  full_year=setNames(full_year,loc_seq[sind])
  print(full_year)
}
```



## Data importing and cleaning

```{r data import}
## Data import
air <- read.csv("tributary air temperatures clean.csv", stringsAsFactors=F)
water <- read.csv("tributary water temperature/water_temperature_d.csv",
                  stringsAsFactors=F)
flow <- read.csv("tributary discharge.csv", stringsAsFactors=F)

# Convert to date format
air$date <- as.Date(air$date, "%m/%d/%Y")
water$date <- as.Date(water$date, "%m/%d/%Y")
flow$date <- as.Date(flow$date, "%m/%d/%Y")


## Calculate the MEAN airT for US locations
for (i in which(is.na(air$mean_temp))) {
  air$mean_temp <- (air$max_temp + air$min_temp) / 2
}
```

### Data combining

```{r data combine}
## Combine water temperature and discharge
aw <- merge(air, water, by = c("station_name","date"))
awf <- merge(aw, flow, by = c("location","date"))
awf <- awf %>% arrange(location, date)

## Change into factor
awf$location <- as.factor(awf$location)
awf$station_name <- as.factor(awf$station_name)

# Get location sequence
loc_seq=levels(awf$location)


## Check if there are any duplicates
duplicates <- awf %>%
  group_by(location, date) %>%
  filter(n() > 1) # Duplicates should be NA...


## Print the result
table(awf$location)
```

### Check for imputed values

Use a 7-days rolling mean to check for possibly imputed temperatures. If the variance of a certain day's temperature is very close to zero, then it is likely that this particular data is imputed.

```{r impute check}
# make sure that no initial NA values in awf water temperature
which(is.na(awf$temp))

# Need to calculate the rolling mean for each location separately...
for(loc in loc_seq) {
  sub <- awf[awf$location == loc,]
  sub$rolling_mean <- rollmean(sub$temp, k = 7, fill = NA, align = "right")
  awf[awf$location == loc, "rolling_mean"] <- sub$rolling_mean
}

# Calculate variance
awf$variance <- (awf$temp - awf$rolling_mean)^2

# Assign NA when variance is very small
awf$temp[which(awf$variance < 1e-10)] <- NA

# Check results - how many imputed values are in each location
awf %>%
  group_by(location) %>%
  summarise(na_count = sum(is.na(temp)))

```


### Get lagged days

```{r lagged days}
## Get the time lag day variables
awf <- awf %>%
  group_by(location) %>%
  mutate(dmean_1 = lag(mean_temp, 1),
         dmean_2 = lag(mean_temp, 2),
         dmean_3 = lag(mean_temp, 3),
         dmean_4 = lag(mean_temp, 4),
         dmean_5 = lag(mean_temp, 5),
         dflow_1 = lag(flow, 1))


## Get cumulative air temp for past five days
awf <- awf %>% 
  rowwise() %>% 
  mutate(cair = (mean_temp+dmean_1+dmean_2+dmean_3+dmean_4+dmean_5)/6)

## Get relative flow
awf <- awf %>% mutate(rqc = (flow - dflow_1)/flow)
```


### Get final master temp dataframe

```{r master temp}
master.temp <- awf[complete.cases(cbind(awf$mean_temp,awf$temp)),] %>% 
  select(location, date, station_name, country,
         year, month = month.x, day = day.x,
         water = temp, air = mean_temp, dmean_1, dmean_2, dmean_3,
         dmean_4, dmean_5, flow, dflow_1, rqc, cair)

master.temp <- master.temp[complete.cases(master.temp[,8:18]),]
```



## Get training and testing


### Subsetting seasonal scales

Now we want to subset three master dataframes that contains seasonal-scale data. We categorize the data into four different seasonal categories:

1. spring: 3,4,5
2. summer: 6,7,8
3. fall: 9,10,11
4. winter: 12,1,2

```{r subset season}
master.sum <- master.temp %>% 
  filter(month == 6 | month == 7 | month == 8)

master.win <- master.temp %>% 
  filter(month == 12 | month == 1 | month == 2)

master.spring <- master.temp %>% 
  filter(month == 3 | month == 4 | month == 5)

master.fall <- master.temp %>% 
  filter(month == 9 | month == 10 | month == 11)

master.annual <- master.temp %>% 
  filter(month != 12 & month != 1 & month != 2)
```


### Identify good year/month data

We want the seasonal data to be greater than 60 days, annual data (winter removed) more than 180 days.

```{r good years}
## Annual (250 days)
fulyear <- good_year(master.annual, 180)


## Seasonal (60 days)
fulspring <- good_year(master.spring, 60)
fulsum <- good_year(master.sum, 60)
fulfall <- good_year(master.fall, 60)
fulwin <- good_year(master.win, 60)
```



### Get training and testing

Similarly, get training and testing for the specific season.

```{r get training and testing}
fold = 10

## Get training and testing for annual
annual <- get.traintest(master.annual, fulyear, 10)

combined_training_list <- annual[[1]]
combined_testing_list <- annual[[2]]


## Get training and testing for each season
sum <- get.traintest(master.sum, fulsum, 10)
win <- get.traintest(master.win, fulwin, 10)
spr <- get.traintest(master.spring, fulspring, 10)
fall <- get.traintest(master.fall, fulfall, 10)

combined_training_list_sp <- spr[[1]]
combined_testing_list_sp <- spr[[2]]

combined_training_list_su <- sum[[1]]
combined_testing_list_su <- sum[[2]]

combined_training_list_fa <- fall[[1]]
combined_testing_list_fa <- fall[[2]]

combined_training_list_w <- win[[1]]
combined_testing_list_w <- win[[2]]


## Create a list to store all the combined training and testing lists
grand_training <- list(combined_training_list_sp, combined_training_list_su,
                       combined_training_list_fa, combined_training_list_w,
                       combined_training_list)

grand_testing <- list(combined_testing_list_sp, combined_testing_list_su,
                       combined_testing_list_fa, combined_testing_list_w,
                       combined_testing_list)
```


## Model Comparison

We now run each of the three models ten times, once on each of the four seasons, and annual data.

**Important: Here we only look at the model output from models with corAR1() value fixed at 0.8. To check the ACF, please refer to the TEMP temporal autocorrelation document.**


### Linear lag5

```{r lag5}
## Forms
ctrl = lmeControl(opt='optim')
form1 <- water ~ air + dmean_1 + dmean_2 + dmean_3 + dmean_4 + dmean_5

coef.lag5 <- data.frame(matrix(NA,50,7))
colnames(coef.lag5) <- c("a","b","c","d","e","f","g")
seasons <- rep(c(1,2,3,4,5), each = 10)
coef.lag5$season <- seasons

spring.r <- vector("list", fold)
summer.r <- vector("list", fold)
fall.r <- vector("list", fold)
winter.r <- vector("list", fold)
annual.r <- vector("list", fold)

grand.lag5 <- list(spring.r, summer.r, fall.r, winter.r, annual.r)


## Iteration starts here
for (season in 1:5) {
  
  ## Get current season and its correspnding training/testing
  train <- grand_training[[season]]
  test <- grand_testing[[season]]
  
  ## 10 fold iteration starts here
  for (i in 1:fold){
    
    # select current dataset, and all unique location levels
    current.training <- train[[i]] %>% arrange(location, date)
    current.testing <- test[[i]] %>% arrange(location, date)
    compare <- current.testing

    # model training and predicting
    model.ar <- lme(form1,
                    random = ~1 | location, control = ctrl,
                    na.action = na.omit, data = current.training,
                    correlation=corAR1(form=~1|location, 0.85, fixed=T))

    compare$preds.ar <- as.vector(predict(
      model.ar, newdata = current.testing, re.form = ~1|location))
    
    grand.lag5[[season]][[i]] <- compare
    coef.lag5[10*(season-1)+i,1:7] = coef(model.ar)[1,]
    #intercept only for bigcreek
  }
}


## Plots
for (loc in loc_seq) {
  df <- grand.lag5[[5]][[1]]
  plot.df <- df[df$location == loc,]
  diff.ar <- round(sqrt(mean((plot.df$preds.ar-plot.df$water)^2)),2)

  pl <- ggplot(data=plot.df, aes(x=date))+
          geom_line(aes(x=date, y=water), color = "black")+
          geom_line(aes(x=date, y=preds.ar), color = "blue")+
          ggtitle(paste(
            loc, ": ", plot.df$year, " (RMSE:", diff.ar, ")"))
  print(pl)
}
```


### Seasonal residual

```{r stochastic}
coef.sto <- data.frame(matrix(NA,50,7))
colnames(coef.sto) <- c("a","b","c","beta1","beta2","beta3","d")
seasons <- rep(c(1,2,3,4,5), each = 10)
coef.sto$season <- seasons


spring.r <- vector("list", fold)
summer.r <- vector("list", fold)
fall.r <- vector("list", fold)
winter.r <- vector("list", fold)
annual.r <- vector("list", fold)

grand.sto <- list(spring.r, summer.r, fall.r, winter.r, annual.r)


## Iteration starts here
for (season in 1:5) {
  
  ## Get current season and its correspnding training/testing
  train <- grand_training[[season]]
  test <- grand_testing[[season]]
  
  # Each iteration
  for (i in 1:fold) {
  
    compare <- NA
    # select current dataset, and all unique location levels
    current.training <- train[[i]] %>% arrange(location, date)
    current.testing <- test[[i]] %>% arrange(location, date)

    ## TRAINING
    # get the model annual component
    annual.comp <- nls(air ~ a+b*sin(2*pi/365*(yday(date)+t0)),
                       start = list(a=0.05, b=5, t0=-26),
                       data=current.training)

    # get the air temperature residuals
    res <- as.data.frame(matrix(NA, ncol = 2, 
                                nrow = length(na.omit(current.training$air))))
    # dataframe to store the residuals
    colnames(res) <- c("res.t", "location")
    res[,"location"] <- na.omit(current.training$location)
    res[,"res.t"] <- as.vector(residuals(annual.comp))
    res <- res %>% group_by(location) %>%
      mutate(res.t1 = lag(res.t, 1), res.t2 = lag(res.t, 2))
    res[,"res.w"] <- residuals(nls(water ~ a+b*sin(2*pi/365*(yday(date)+t0)),
                               start = list(a=0.05, b=5, t0=-26),
                               data = current.training))
    
    # get the water temperature residual component
    residual.comp.ar <- lme(fixed = res.w ~ res.t + res.t1 + res.t2,
                            random = ~ 1|location,
                            correlation = corAR1(form=~1|location,
                                                 0.85,fixed=T),
                            data = res, na.action = na.omit,
                            control = ctrl)
  
    ## TESTING
    # Annual
    preds.annual <- as.data.frame(
      predict(annual.comp,newdata=current.testing))
    preds.annual <- cbind(preds.annual,
                          current.testing$location, current.testing$date)
    colnames(preds.annual) <- c("preds.annual", "location", "date")

    # Residuals
    res <- as.data.frame(matrix(NA, ncol = 3, 
                                nrow=length(current.testing$air))) #residuals
    colnames(res) <- c("res.t", "location", "date")
    res[,"location"] <- current.testing$location
    res[,"date"] <- current.testing$date
    res[,"res.t"] <- current.testing$air - preds.annual$preds.annual
    res <- res %>% group_by(location) %>%
      mutate(res.t1 = lag(res.t, 1),res.t2 = lag(res.t, 2))
    
    pres.ar <- predict(residual.comp.ar, newdata=res, na.action=na.omit,
                       re.form=~(1|location))
    preds.residuals <- cbind(na.omit(res)[,"location"],
                             na.omit(res)[,"date"],as.data.frame(pres.ar))
  
    # add up both components
    p <- merge(preds.annual, preds.residuals, by=c("location","date"))
    p[,"preds.ar"] <- p$preds.annual + p$pres.ar

    ## Calculate RMSE
    compare <- merge(current.testing, p, by=c("location","date")) %>% 
      select(location, year, date, water, preds.ar, preds.annual)
    
    grand.sto[[season]][[i]] <- compare
    coef.sto[10*(season-1)+i,1:3] = coef(annual.comp)
    coef.sto[10*(season-1)+i,4:6] = coef(residual.comp.ar)[1,2:4]
    coef.sto[10*(season-1)+i,7] = coef(residual.comp.ar)[1,1]
  }
}


## Plots
for (loc in loc_seq) {
  df <- grand.sto[[5]][[1]]
  plot.df <- df[df$location == loc,]
  diff.ar <- round(sqrt(mean((plot.df$preds.ar-plot.df$water)^2)),2)

  pl <- ggplot(data=plot.df, aes(x=date))+
          geom_line(aes(x=date, y=water), color = "black")+
          geom_line(aes(x=date, y=preds.ar), color = "blue")+
          ggtitle(paste(
            loc, ": ", plot.df$year, " (RMSE:", diff.ar, ")"))
  print(pl)
}
```


### Non-linear

```{r non-linear, eval=TRUE, warning=FALSE}
coef.nonlinear <- data.frame(matrix(NA,50,3))
colnames(coef.nonlinear) <- c("alpha","gamma","beta")
seasons <- rep(c(1,2,3,4,5), each = 10)
coef.nonlinear$season <- seasons


## starting parameters
coef.spring <- c(alpha=30, gamma=0.05,beta=10)
coef.summer <- c(alpha=25, gamma=0.05,beta=10)
coef.fall <- c(alpha=30, gamma=0.05,beta=10)
coef.annual <- c(alpha=20, gamma=0.05,beta=9)

coef.list <- list(coef.spring, coef.summer, coef.fall, NA, coef.annual)


spring.r <- vector("list", fold)
summer.r <- vector("list", fold)
fall.r <- vector("list", fold)
winter.r <- vector("list", fold)
annual.r <- vector("list", fold)

grand.nonlinear <- list(spring.r, summer.r, fall.r, winter.r, annual.r)


## Iteration starts here
for (season in c(1,2,3,5)) {
  
  ## Get current season and its correspnding training/testing
  train <- grand_training[[season]]
  test <- grand_testing[[season]]
  
  ## 10 fold iteration starts here
  for (i in 1:fold){
  
    compare <- NA
    # select current dataset, and all unique location levels
    current.training <- train[[i]]
    current.testing <- test[[i]]
    compare <- current.testing

    # model training and predicting
    model <- nlme(water ~ alpha / (1 + exp(gamma * (beta - cair))),
                  fixed = list(alpha~1,gamma~1,beta~1),
                  random = gamma ~ 1|location,
                  start = coef.list[[season]],
                  data = current.training, na.action = na.omit,
                  control = list(msMaxIter = 200))
    
    compare$preds.ar <- as.vector(predict(
      model, newdata = current.testing, re.form = ~1|location))
  
    grand.nonlinear[[season]][[i]] <- compare
    coef.nonlinear[10*(season-1)+i,1:3] <- coef(model)
  }
}
```


### lag5 with flow

```{r flow}
## Forms
ctrl = lmeControl(opt='optim')
form3 <- water ~ air + dmean_1 + dmean_2 + dmean_3 + dmean_4 + dmean_5 + I(1/flow)

spring.r <- vector("list", fold)
summer.r <- vector("list", fold)
fall.r <- vector("list", fold)
winter.r <- vector("list", fold)
annual.r <- vector("list", fold)

grand.flow <- list(spring.r, summer.r, fall.r, winter.r, annual.r)


## Iteration starts here
for (season in 1:5) {
  
  ## Get current season and its correspnding training/testing
  train <- grand_training[[season]]
  test <- grand_testing[[season]]
  
  ## 10 fold iteration starts here
  for (i in 1:fold){
    
    # select current dataset, and all unique location levels
    current.training <- train[[i]] %>% arrange(location, date)
    current.testing <- test[[i]] %>% arrange(location, date)
    compare <- current.testing

    # model training and predicting
    model.ar <- lme(form3,
                    random = ~1 | location, control = ctrl,
                    na.action = na.omit, data = current.training,
                    correlation=corAR1(form=~1|location, 0.85, fixed=T))

    compare$preds.ar <- as.vector(predict(
      model.ar, newdata = current.testing, re.form = ~1|location))
    
    grand.flow[[season]][[i]] <- compare
  }
}


## Plots
for (loc in loc_seq) {
  df <- grand.flow[[5]][[1]]
  plot.df <- df[df$location == loc,]
  diff.ar <- round(sqrt(mean((plot.df$preds.ar-plot.df$water)^2)),2)

  pl <- ggplot(data=plot.df, aes(x=date))+
          geom_line(aes(x=date, y=water), color = "black")+
          geom_line(aes(x=date, y=preds.ar), color = "blue")+
          ggtitle(paste(
            loc, ": ", plot.df$year, " (RMSE:", diff.ar, ")"))
  print(pl)
}
```


### Comparison at the end

We first compare the RMSE of each model within one specific season.

```{r comparison model by season}
## Spring::
spring.compare <- data.frame(
  lag5 = colMeans(cal.rmse(grand.lag5[[1]], fold)),
  sto = colMeans(cal.rmse(grand.sto[[1]], fold)),
  nonlinear = colMeans(cal.rmse(grand.nonlinear[[1]], fold)),
  flow = colMeans(cal.rmse(grand.flow[[1]], fold)))

colMeans(spring.compare)
knitr::kable(spring.compare, digits = 3)


## Summer::
summer.compare <- data.frame(
  lag5 = colMeans(cal.rmse(grand.lag5[[2]], fold), na.rm=T),
  sto = colMeans(cal.rmse(grand.sto[[2]], fold), na.rm=T),
  nonlinear = colMeans(cal.rmse(grand.nonlinear[[2]], fold), na.rm=T),
  flow = colMeans(cal.rmse(grand.flow[[2]], fold), na.rm=T))

colMeans(summer.compare)
knitr::kable(summer.compare, digits = 3)


## Fall::
fall.compare <- data.frame(
  lag5 = colMeans(cal.rmse(grand.lag5[[3]], fold)),
  sto = colMeans(cal.rmse(grand.sto[[3]], fold)),
  nonlinear = colMeans(cal.rmse(grand.nonlinear[[3]], fold)),
  flow = colMeans(cal.rmse(grand.flow[[3]], fold)))

colMeans(fall.compare)
knitr::kable(fall.compare, digits = 3)


## Annual
annual.compare <- data.frame(
  lag5 = colMeans(cal.rmse(grand.lag5[[5]], fold)),
  sto = colMeans(cal.rmse(grand.sto[[5]], fold)),
  nonlinear = colMeans(cal.rmse(grand.nonlinear[[5]], fold)),
  flow = colMeans(cal.rmse(grand.flow[[5]], fold)))

colMeans(annual.compare)
knitr::kable(annual.compare, digits = 3)
```

Now, let's compare the NSC of each model within one specific season.

```{r comparison model by season}
## Spring::
spring.compare <- data.frame(
  lag5 = colMeans(cal.nsc(grand.lag5[[1]], fold)),
  sto = colMeans(cal.nsc(grand.sto[[1]], fold)),
  nonlinear = colMeans(cal.nsc(grand.nonlinear[[1]], fold)),
  flow = colMeans(cal.nsc(grand.flow[[1]], fold)))

colMeans(spring.compare)
knitr::kable(spring.compare, digits = 3)


## Summer::
summer.compare <- data.frame(
  lag5 = colMeans(cal.nsc(grand.lag5[[2]], fold), na.rm=T),
  sto = colMeans(cal.nsc(grand.sto[[2]], fold), na.rm=T),
  nonlinear = colMeans(cal.nsc(grand.nonlinear[[2]], fold), na.rm=T),
  flow = colMeans(cal.nsc(grand.flow[[2]], fold), na.rm=T))

colMeans(summer.compare)
knitr::kable(summer.compare, digits = 3)


## Fall::
fall.compare <- data.frame(
  lag5 = colMeans(cal.nsc(grand.lag5[[3]], fold)),
  sto = colMeans(cal.nsc(grand.sto[[3]], fold)),
  nonlinear = colMeans(cal.nsc(grand.nonlinear[[3]], fold)),
  flow = colMeans(cal.nsc(grand.flow[[3]], fold)))

colMeans(fall.compare)
knitr::kable(fall.compare, digits = 3)


## Annual
annual.compare <- data.frame(
  lag5 = colMeans(cal.nsc(grand.lag5[[5]], fold)),
  sto = colMeans(cal.nsc(grand.sto[[5]], fold)),
  nonlinear = colMeans(cal.nsc(grand.nonlinear[[5]], fold)),
  flow = colMeans(cal.nsc(grand.flow[[5]], fold)))

colMeans(annual.compare)
knitr::kable(annual.compare, digits = 3)
```



We then compare the results of different seasonal scales on each individual model.

```{r comparison season by model, eval=FALSE}
## Lag5::
lag5.compare <- data.frame(
  spring = colMeans(cal.rmse(grand.lag5[[1]], fold)),
  summer = colMeans(cal.rmse(grand.lag5[[2]], fold), na.rm=T),
  fall = colMeans(cal.rmse(grand.lag5[[3]], fold)),
  annual = colMeans(cal.rmse(grand.lag5[[5]], fold)))

colMeans(lag5.compare)
knitr::kable(lag5.compare, digits = 3)


## Stochastic::
sto.compare <- data.frame(
  spring = colMeans(cal.rmse(grand.sto[[1]], fold)),
  summer = colMeans(cal.rmse(grand.sto[[2]], fold), na.rm=T),
  fall = colMeans(cal.rmse(grand.sto[[3]], fold)),
  annual = colMeans(cal.rmse(grand.sto[[5]], fold)))

colMeans(sto.compare)
knitr::kable(sto.compare, digits = 3)


## Flow::
nonlinear.compare <- data.frame(
  spring = colMeans(cal.rmse(grand.nonlinear[[1]], fold)),
  summer = colMeans(cal.rmse(grand.nonlinear[[2]], fold),na.rm=T),
  fall = colMeans(cal.rmse(grand.nonlinear[[3]], fold)),
  annual = colMeans(cal.rmse(grand.nonlinear[[5]], fold)))

colMeans(nonlinear.compare)
knitr::kable(flow.compare, digits = 3)
```

### Extract model coefficients

```{r coefs}
View(coef.lag5)

for(s in 1:5) {
  c = coef.lag5[coef.lag5$season == s,]
  print(round(colMeans(c),2))
}


```


